--- 
title: "A Minimal Book Example"
author: "Yihui Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---
```{r include=FALSE}
# not run any of the code chunks when publish
knitr::opts_chunk$set(eval = F)
```

# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

<!--chapter:end:index.Rmd-->


# Load data
```{r include=FALSE}
# not run any of the code chunks when publish
knitr::opts_chunk$set(eval = F)
```

## csv
```{r eval=FALSE}
library(data.table)
infile <- 'input_file.csv'
df <- fread(infile, data.table = F)

```

## excel
### general excel
```{r}
library(xlsx)

# simply load
df <- xlsx::read.xlsx(infile, sheetIndex = 1)
df <- xlsx::read.xlsx(infile, sheetName = 'SheetName')

# row/col customized
df <- xlsx::read.xlsx(infile,
                      sheetIndex = 1,
                      startRow = 2,
                      colIndex = excelCol2Num[c('AA','B')], # helper vector below
                      endRow = 99)
```

### excel with password
```{r}
library(xlsx)
df <- read.xlsx(file.xlsx,
                sheetIndex = 1,
                password = 'password')

```


### larger excel
```{r}
library(openxlsx)
df <- openxlsx::read.xlsx(infile)
```


### excel column letter to index number
```{r}
excelCol2Num <- 1:702
names(excelCol2Num) <- do.call(paste0, expand.grid(LETTERS, c("", LETTERS))[,2:1])

excelCol2Num[c('AA', 'B')]
```

### header name keep special character
```{r}
library(janitor)
df_raw <- read.csv(infile, check.names = F)
df <- clean_names(df_raw)

ref_colnames <- data.frame(original = colnames(df_raw), 
                           row.names = make_clean_names(colnames(df_raw)))


```

## txt vector
```{r}
v1 <- scan('file.txt', what = character, sep = '\n')
```


<!--chapter:end:01-load.Rmd-->



# Explore Data


## columns quick view
```{r eval=FALSE}
str(df)

apply(df, 2, function(x) range(x, na.rm = T))

```


<!--chapter:end:02-exploreTable.Rmd-->

```{r include=FALSE}
# not run any of the code chunks when publish
knitr::opts_chunk$set(eval = F)
```


# clean data



## duplicates
```{r}
df <- df[ !duplicated(df), ]
```

## column types
```{r}
library(dplyr)
df <- df %>% mutate_all(as.character)

```



## string manipulation
### replace
```{r}
df$text <- gsub('old', 'new', df$text)
```

### split
```{r}
text_part1 <- sapply(strsplit(df$text_full, ' '), '[', 1)
```



## merge
```{r}
df <- merge(df1, df2, by=c('id'), all.x = T)

df <- df1 %>%
  left_join(., df2, by = c('id'))
```

## select column by name
```{r}
df <- subset(df, select = c(id, gender))

df <- df %>%
  select('id', 'age')
```


<!--chapter:end:03-clean.Rmd-->

# subset

Some _significant_ applications are demonstrated in this chapter.

## Example one

## Example two

<!--chapter:end:04-subset.Rmd-->

# output

## csv
```{r}
write.csv(df, row.names = F)
```


<!--chapter:end:05-writeTable.Rmd-->

`r if (knitr::is_html_output()) '
# useful links {-}
'`

<!--chapter:end:06-plotHeatmap.Rmd-->

